Concurrent and Distributed Programming in Harmony

### 2022-01-11 paper/paper.tex
Latest version of paper

<{:const:tas}>
<{:const:cas}>
<{:const:BinSem}>
<{:const:Semaphore}>
<{:const:__init__}>
<{:const:p}>
<{:const:hello}>
<{:const:enqueue}>
<{:const:dequeue}>

@@

@@

@@

for i in {0..N-1}: spawn diner(i)
<{:end:}>
\end{code}
}
\end{center}
\caption{The Dining Philosophers Program in Harmony}
\label{fig:diners}
\end{figure}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{diners.png}
\end{center}
\caption{The output of running the Dining Philosophers program}
\label{fig:dinersHTML}
\end{figure*}

@@

However, users first want to focus on the rows in the top table.
In this particular example

>>>>

<{:const:tas}>
<{:const:cas}>
<{:const:BinSem}>
<{:const:Semaphore}>
<{:const:__init__}>
<{:const:p}>
<{:const:hello}>
<{:const:enqueue}>
<{:const:dequeue}>
<{:const:send}>
<{:const:receive}>

@@

This paper provides an overview of the Harmony language and Charm,
its model checker, and how they may be used by programmers to
develop reliable concurrent or distributed code.

@@

\subsection{Distributed Programming}

Harmony can be used to model a distributed system in which the
only shared variable is a set of messages representing the network.
A message is sent by atomically adding a message to the network:
<{def send(m): atomically network |= {m}}>.
To receive a message, a program can use the Harmony <{when}>
statement to wait for a message.  Typically, however, the program
does not want to receive just any message, but a message that satisfies
certain conditions such as: wait for the next message that is addressed
to me and has a sequence number larger than 3.
For this, Harmony supports the <{exists clause}> and allows
<{when}> clauses to be nested:
<{atomically when exists m in network when (m.dst == self) and (m.seq > 3): ...}>
With these constructs, Harmony can be effectively used as a
model checker for distributed algorithms.

@@

for i in {0..N-1}: spawn diner(i)
<{:end:}>
\end{code}
}
\end{center}
\caption{Dining Philosophers in Harmony}
\label{fig:diners}
\end{figure}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{diners.png}
\end{center}
\caption{The HTML output of running the Dining Philosophers program}
\label{fig:dinersHTML}
\end{figure*}

@@

However, users first want to focus on the rows in the top table.
In this particular example, there are 6 threads, the thread that
first runs and the five threads that are spawned by it.  They take
``turns'' by context switching.  In this particular trace, <{diner(2)}>
runs first and acquires <{forks[2]}>, its left fork.  In the table
in the lower-right, you can see that the thread is currently blocked
and in atomic as well as read-only mode (conditions in a <{when}>
clause are not allowed to modify state).  You can also see its stack
trace: it trying to acquire <{forks[3]}>, its right fork (executing
the code in \autoref{fig:synch}(a)).  Next <{diner(0)}>, <{diner(3)}>,
<{diner(4)}>, and <{diner(1)}> take similar turns.  All threads are
blocked waiting for another thread to release a lock---a classic
deadlock situation.

Should the programmer need more detailed information, then they can
run through the counter-example step-by-step, forwards and backwards,
as explained above.

\section{Limitations}

Combinatorial state explosion.
Production code.
Lack of formal definition of Harmony itself.
